---
title: "part 2 slides"
author: "Xi Cheng"
date: "March 20, 2018"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(stringr)
library(ggplot2)
library(RColorBrewer)
```
## Cleaning the original dataset

```{r,message=FALSE,Warnings= FALSE}
options(warn=-1)
veg.1 <- read_xlsx("veg1.xlsx")


# First use code from Class 19

cnames.1 <- colnames(veg.1)

c <- apply(veg.1, 2, n_distinct)

c[c>1]

d <- names(c[c==1])

e <- names(c[c>1])

# delete all useless columns (which only have one distinct value)

veg.2 <- select(veg.1, e)

cnames.2 <- colnames(veg.2)

apply(veg.2, 2, n_distinct)

# change some columns names 
veg.3 <- dplyr::rename(veg.2, 
                       Geo = `Geo Level`, 
                       State = `State ANSI`,
                       Data = `Data Item`,
                       Category = `Domain Category`)

cnames.3 <- colnames(veg.3)

a1 <- separate(veg.3, Category, into = c("Label", "Quant"), sep=",")

# We only focus on the data with "RESTRICTED USE CHEMICAL"

a2 <- filter(a1, Label=="RESTRICTED USE CHEMICAL")

# We have 2 commodity left, BROCCOLI and CAULIFLOWER

unique(a2[,"Commodity"])

a21 <- a2 %>% select(Label, Quant) %>% unique()

# We have about 28 chemicals left 

a21 %>% print(n=30)

# Keep clean the dataset 

a3 <- separate(a2, Quant, into=c("Treatment Product","Chemical and Code"), sep = ":")

# Delete unnecessary column

a4 <- select(a3, -Domain)

a5 <- separate(a4 ,Data, into = c("Commodity Name", "Measureed Value"), sep = "-")

a6 <- separate(a5, "Measureed Value", into = c("Application or Treated", "Way of Measurement", "AVG or not"), sep = ",")

unique(a6[,"Geo"])

unique(a6[,"State"])

unique(a6[,"Region"])

# Delete dupilcate column and columns that has only one distinct value

a7 <- select(a6, -Geo, -State, -Region, -Commodity)

a8 <- a7 %>% separate(`Chemical and Code` , into = c("Chemical Name", "Code"), sep = "=") %>% 
  separate(`Chemical Name`, into = c("Useless1", "Chemical Name"), sep = "[()]") %>%
  separate(Code, into = c("Chemical Code", "Useless2"), sep = "[())]") %>%
  select(-Useless1, -Useless2)

# For different ways of Measurement, we are only instersted in the value of each chemical "MEASURED IN LB"

a9 <- filter(a8, `Way of Measurement`==" MEASURED IN LB")
             
# Delete the chemicals whose value we don't know

a10 <- filter(a9, !Value %in% c("(D)", "(Z)", "(NA)", NA))

# Delete unnecessary column

a11 <- select(a10, -`AVG or not`)

a11$Value <- as.numeric(a11$Value)
a11$Year <- as.character(a11$Year)
a11$`Chemical Name` <- as.character(a11$`Chemical Name`)

a11

```
##Dealing with LD5O data for each chemicals 
we just want to find out the LD50 data for each chenicals, so that we could do more things about all these chemicals.From which, we could find out that the DISULFOTON is the most toxical one, and the PRONAMIDE is the least one.
```{r}

# Now, we only need to check the LD50 for 15 chemicals
# They are:
# 1 "BETA-CYFLUTHRIN "    
# 2 "BIFENTHRIN "         
# 3 "ESFENVALERATE "      
# 4 "IMIDACLOPRID "       
# 5 "LAMBDA-CYHALOTHRIN " 
# 6 "METHOMYL "           
# 7 "NALED "              
# 8 "PERMETHRIN "         
# 9 "ZETA-CYPERMETHRIN "  
# 10 "CHLORANTRANILIPROLE "
# 11 "CHLORPYRIFOS "       
# 12 "DIAZINON "           
# 13 "PRONAMIDE "          
# 14 "DISULFOTON "         
# 15 "EMAMECTIN BENZOATE " 

# Except BETA-CYFLUTHRIN and ZETA-CYPERMETHRIN, 
# we can find the LD50 for all other above chemicals from https://comptox.epa.gov/dashboard
# Let's build a table for the LD50 of the other 13 chemicals
# Then we can add the LD50 we find from other website to this table

BIFENTHRIN <- read.csv("BIFENTHRIN.csv")
as.tibble(BIFENTHRIN)
ESFENVALERATE <- read.csv("ESFENVALERATE.csv")
as.tibble(ESFENVALERATE)
IMIDACLOPRID <- read.csv("IMIDACLOPRID.csv")
as.tibble(IMIDACLOPRID)
LAMBDA_CYHALOTHRIN <- read.csv("LAMBDA-CYHALOTHRIN.csv")
as.tibble(LAMBDA_CYHALOTHRIN)
METHOMYL <- read.csv("METHOMYL.csv")
as.tibble(METHOMYL)
NALED <- read.csv("NALED.csv")
as.tibble(NALED)
PERMETHRIN <- read.csv("PERMETHRIN.csv")
as.tibble(PERMETHRIN)
CHLORANTRANILIPROLE <- read.csv("CHLORANTRANILIPROLE.csv")
as.tibble(CHLORANTRANILIPROLE)
CHLORPYRIFOS <- read.csv("CHLORPYRIFOS.csv")
as.tibble(CHLORPYRIFOS)
DIAZINON <- read.csv("DIAZINON.csv")
as.tibble(DIAZINON)
PRONAMIDE <- read.csv("PRONAMIDE.csv")
as.tibble(PRONAMIDE)
DISULFOTON <- read.csv("DISULFOTON.csv")
as.tibble(DISULFOTON)
EMAMECTIN_BENZOATE <- read.csv("EMAMECTIN BENZOATE.csv")
as.tibble(EMAMECTIN_BENZOATE)

# The experiment animal we choose is rat

BIFENTHRIN <- BIFENTHRIN %>% mutate( `Chemical Name` = "BIFENTHRIN") %>% filter(SPECIES == "rat")
ESFENVALERATE <- ESFENVALERATE %>% mutate( `Chemical Name` = "ESFENVALERATE") %>% filter(SPECIES == "rat")
IMIDACLOPRID <- IMIDACLOPRID %>% mutate( `Chemical Name` = "IMIDACLOPRID") %>% filter(SPECIES == "rat")
LAMBDA_CYHALOTHRIN <- LAMBDA_CYHALOTHRIN %>% mutate( `Chemical Name` = "LAMBDA-CYHALOTHRIN") %>% filter(SPECIES == "rat")
METHOMYL <- METHOMYL %>% mutate( `Chemical Name` = "METHOMYL") %>% filter(SPECIES == "rat")
NALED <- NALED %>% mutate( `Chemical Name` = "NALED") %>% filter(SPECIES == "rat")
PERMETHRIN <- PERMETHRIN %>% mutate( `Chemical Name` = "PERMETHRIN") %>% filter(SPECIES == "rat")
CHLORANTRANILIPROLE <- CHLORANTRANILIPROLE %>% mutate( `Chemical Name` = "CHLORANTRANILIPROLE") %>% filter(SPECIES == "rat")
CHLORPYRIFOS <- CHLORPYRIFOS %>% mutate( `Chemical Name` = "CHLORPYRIFOS") %>% filter(SPECIES == "rat")
DIAZINON <- DIAZINON %>% mutate( `Chemical Name` = "DIAZINON") %>% filter(SPECIES == "rat")
PRONAMIDE <- PRONAMIDE %>% mutate( `Chemical Name` = "PRONAMIDE") %>% filter(SPECIES == "rat")
DISULFOTON <- DISULFOTON %>% mutate( `Chemical Name` = "DISULFOTON") %>% filter(SPECIES == "rat")
EMAMECTIN_BENZOATE <- EMAMECTIN_BENZOATE %>% mutate( `Chemical Name` = "EMAMECTIN BENZOATE") %>% filter(SPECIES == "rat")


# Combine them into one table, and the next step is to clean it

LD50_1 <- rbind(BIFENTHRIN, ESFENVALERATE, IMIDACLOPRID, LAMBDA_CYHALOTHRIN,
              METHOMYL, NALED, PERMETHRIN, CHLORANTRANILIPROLE, CHLORPYRIFOS, 
              DIAZINON, PRONAMIDE, DISULFOTON, EMAMECTIN_BENZOATE)


unique(LD50_1[,"Chemical Name"])

LD50_2 <- filter(LD50_1, UNITS  == "mg/kg")

unique(LD50_2[,"SOURCE"])

LD50_3 <- filter(LD50_2, SOURCE  == "AcuteTox")

unique(LD50_3[,"Chemical Name"])

unique(LD50_3[,"TYPE"])

unique(LD50_3[,"STUDY_TYPE"])


# Now we have make sure each sample is oral LD50 measured on rats in mg/kg from source AcuteTox
# But we don't have this data for EMAMECTIN BENZOATE after filtering, 
# so besides BETA-CYFLUTHRIN and ZETA-CYPERMETHRIN, we also need to find 
# the data of EMAMECTIN_BENZOATE from other sources. 
# Before that, let's keep cleaning the data

LD50_4 <- LD50_3 %>% group_by(`Chemical Name`) %>%
  dplyr::summarize(round(mean(VALUES)))

LD50_4 <- plyr::rename(LD50_4, c(`round(mean(VALUES))` = "LD50 for Rats (mg/kg)" ))

# From http://pmep.cce.cornell.edu, we get the LD5O on rats for BETA-CYFLUTHRIN and ZETA-CYPERMETHRIN,
# are 869 - 1271 mg/kg(take average as 1070 mg/kg) and 250 mg/kg respectively. 
# From http://www.rayfull.com/UploadFiles/PDF/201368165633.pdf, we know that the LD50 on rats for 
# EMAMECTIN BENZOATE is 1516 mg/kg. 
# Add these into LD50_4


aa <- c("BETA-CYFLUTHRIN", 1070)
bb <- c("ZETA-CYPERMETHRIN", 250)
cc <- c("EMAMECTIN BENZOATE", 1516)

LD50_4[nrow(LD50_4) + 1,] <- aa
LD50_4[nrow(LD50_4) + 1,] <- bb
LD50_4[nrow(LD50_4) + 1,] <- cc

LD50_4$`LD50 for Rats (mg/kg)` <- as.numeric(options(warn=-1))

LD50_4



```
##Combine tables from Step 1 and 2 as one 
We use what we have gotten in the part 1 and part 2, combine them together so we could get the the combined table tell us about all the maseeges what we want to get. From the table we could see the chemical code, the value of the chemical apllied to commodity, and the LD50. Shows as the table below.
```{r}
# combine two tables by the name of chemicals
# make the chemical name column in a11 is the same in LD50_4

a12 <- a11 %>% separate(`Chemical Name` , into = c("Chemical Name", "nothing"), sep = " ") %>%
  select(-nothing)

a12[46,7] <- "EMAMECTIN BENZOATE"

unique(a12[,"Chemical Name"])
unique(LD50_4[,"Chemical Name"])

merge1 <- left_join(a12, LD50_4, by = "Chemical Name")

# Now we have combined them in to one, but we can make it look better 

merge2 <- merge1[,c(1,2,6:10)]

merge2 <- plyr::rename(merge2, c(`Chemical Name` = "Restricted Use Chemical Name" ))
merge2 <- plyr::rename(merge2, c(Value = "Value of the Chemical Applied to the Commodity in (LB)" ))
merge2 <- plyr::rename(merge2, c(`LD50 for Rats (mg/kg)` = "Chemical's LD50 for Rats (mg/kg)" ))

# We also notice that the unit of Chemical Value and Chemical LD50 are different.
# We also want to more derictly show the effect of chemical toxicity.
# So we unify the unit and add a column to show how many rats could be killed by each amount of chemical 

improve1 <- merge2 %>% mutate("Value of the Chemical Applied to the Commodity in (mg)" = `Value of the Chemical Applied to the Commodity in (LB)` * 453592)

improve1 <- improve1[,c(1:6,8,7)]

# We want to know how many rats can be killed be each amount of chemical 
# We assume the mass of each rat is 300g (0.3kg)
# Prevent Scientific Notation of large numbers
options(scipen=999)
improve2 <- improve1 %>% 
  mutate(round("Number of 300g Rats Can Be Killed" = `Value of the Chemical Applied to the Commodity in (mg)` /(2 * 0.3 * `Chemical's LD50 for Rats (mg/kg)`)))

improve2 <- plyr::rename(improve2, c( "round(`Number of 300g Rats Can Be Killed` = `Value of the Chemical Applied to the Commodity in (mg)`/(2 * \n    0.3 * `Chemical's LD50 for Rats (mg/kg)`))" = "Number of 300g Rats Can Be Killed"))

# Now we have our final cleaned table 
Final_Table <- improve2
Final_Table



```
##Visualization
##We notice that CHLORANTRANILIPROLE and PRONAMIDE has the most toxicity
```{r}
ggplot(LD50_4,aes(`Chemical Name`,`LD50 for Rats (mg/kg)`))+
  geom_bar(stat="identity",fill="#009E73")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "The LD50 on Rats for each Chemical") 
```
##Barplot for Restricted Use Chemical applied on Broccoli by Year
```{r}
part1 <- filter(Final_Table, `Commodity Name`  == "BROCCOLI ")
ggplot(part1, aes( x = `Restricted Use Chemical Name`, y = `Value of the Chemical Applied to the Commodity in (LB)`, fill = Year))+
         geom_bar(position = "dodge",stat="identity") + 
          theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
          scale_fill_brewer(palette = "Greens") +
        labs(title = "Restricted Use Chemical applied on Broccoli by Year")
```
##Barplot for Restricted Use Chemical applied on Cauliflower by Year
```{r}
part2 <- filter(Final_Table, `Commodity Name`  == "CAULIFLOWER ")
ggplot(part2, aes( x = `Restricted Use Chemical Name`, y = `Value of the Chemical Applied to the Commodity in (LB)`, fill = Year))+
  geom_bar(position = "dodge",stat="identity") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  scale_fill_brewer(palette = "Greens") +
  labs(title = "Restricted Use Chemical applied on Cauliflower by Year")
```
##Barplots that shows the Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Broccoli in different years
```{r}
part3 <- filter(Final_Table, `Commodity Name`  == "BROCCOLI ", Year == 2016)
ggplot(part3, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Broccoli in 2016")

part4 <- filter(Final_Table, `Commodity Name`  == "BROCCOLI ", Year == 2014)
ggplot(part4, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Broccoli in 2014")


part5 <- filter(Final_Table, `Commodity Name`  == "BROCCOLI ", Year == 2010)
ggplot(part5, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Broccoli in 2010")

part6 <- filter(Final_Table, `Commodity Name`  == "BROCCOLI ", Year == 2006)
ggplot(part6, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Broccoli in 2006")

```
##Barplots that shows the Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Cauliflower in different years
```{r}
part7 <- filter(Final_Table, `Commodity Name`  == "CAULIFLOWER ", Year == 2016)
ggplot(part7, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Cauliflower in 2016")

part8 <- filter(Final_Table, `Commodity Name`  == "CAULIFLOWER ", Year == 2014)
ggplot(part8, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Cauliflower in 2014")


part9 <- filter(Final_Table, `Commodity Name`  == "CAULIFLOWER ", Year == 2010)
ggplot(part9, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Cauliflower in 2010")

part10 <- filter(Final_Table, `Commodity Name`  == "CAULIFLOWER ", Year == 2006)
ggplot(part10, aes( x = `Restricted Use Chemical Name`, y = `Number of 300g Rats Can Be Killed`))+
  geom_bar(stat="identity", fill = "skyblue") + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  labs(title = "Number of 300g Rats Can Be Killed by Restricted Use Chemical applied on Cauliflower in 2006")



part11 <- filter(Final_Table, `Treatment Product`  == " HERBICIDE")




part11 <- filter(a3, `Treatment Product`  == " HERBICIDE")


```













